---
title: "Problem Set 1 Key"
output: pdf_document
---


**1. Open the dataset as a dataframe. This dataframe should have the following properties: a) The column names should match the column names in the original dataset. b) The row names should correspond to the variable ID in the original dataset.**


```{r, echo=T}
Expends2002 <- read.csv(
"C:/Users/dl0ck/OneDrive/Spring2020/Political Data Science/Problem Sets/PS 1/Expends2002.csv", 
na.strings="", stringsAsFactors=FALSE)
Expends2002 <- as.data.frame(Expends2002)
```


**2. Change the variable name TransID to Useless.**


```{r}

colnames(Expends2002)
colnames(Expends2002)[3] <- "Useless"
colnames(Expends2002)
```

**3. Remove the variables Useless, and Source from the dataframe**

```{r}
Expends2002 <- Expends2002[,-3]
colnames(Expends2002)
```


**4. Change the variable EntType to a factor. How many levels does this variable have?**

```{r}
Expends2002$EntType <- as.factor(Expends2002$EntType)
```

**5. The variable State contains several obvious errors, as it includes nonexistent state codes.**

  + Identify observations that have non-existent state codes.
  
  + Write a script to recode these observations. Use the additional information in the dataset (candidate name, city, zip code) to correctly identify each state.
  
```{r}
unique(Expends2002$State)

wrongstate<-subset(Expends2002,State=="St" | State=="ZZ" | State=="LL")
Expends2002$State[Expends2002$State=="LL"] <- "IA"
Expends2002$State[Expends2002$State=="ZZ"] <- "VI"
Expends2002$State[Expends2002$State=="St"] <- "DC"
```  
  
**6.  Remove all observations from the dataset where the variable State is missing. Report the number of observations after removing missing values.**

```{r}
Expends2002<-subset(Expends2002,State!="  ")
nrow(Expends2002)

```

**7.  Change the variable Zip into a numeric. Be sure to document what you do with missing cases. What is the mean of this variable?**
```{r}
Expends2002$Zip <- as.numeric(Expends2002$Zip)
#Missing cases become NA
mean(Expends2002$Zip, na.rm=T)

```

**8. Create new variables that contain the following information (you will be making several variables), and answer the questions:**


    

    + The number of words in the Descrip variable. What is the median value of this new variable?
```{r, echo=T}
#Part 1
df <- NULL
df$descrip<- Expends2002$Descrip
df <- as.data.frame(df)
df$total <- sapply(df$descrip, function(x) length(unlist(strsplit(as.character(x), "\\W+"))))
sum(df$total, na.rm =T) #The number of words in the Descrip variable
median(df$total, na.rm=T) # The median number of words in each row
```

    + A variable containing the numeric portion of CRPFilerid. This variable should be of length 8 
    for all observations. What is the number of unique values of this variable?
```{r, echo=T}
#Part 2
df$filerid <- Expends2002$CRPFilerid
df$filerid <- as.character(df$filerid)
df$filerid <- sapply(df$filerid, function(x) gsub("[^0-9.-]", "", x)) #get rid of all the leters
length(unique(df$filerid)) #number of unique values in this variable
```

    
    + A vector containing the first four digits of Zip. What is the most frequent value of this vector?
    
```{r, echo =T}
#Part 3
df$zip <- substr(as.character(Expends2002$Zip), 0,4)
#df$zip
names(which(max(table(df$zip))==table(df$zip)))
````

    + A boolean indicating whether the Descrip variable contains the word “Communications” 
    REGARDLESS OF CAPITALIZATION. Report the number of TRUE values in this boolean.
    
```{r, echo =T}
#Part 4
df$descrip <- as.character(df$descrip)
#df$descrip
#grepl("communications", df$descrip, ignore.case=T)
sum(grepl("communications", df$descrip, ignore.case=T)==T)
```

    + A variable indicating that either CRPFilerid is “N” or that BOTH  Amount is greater than 
    500 and Descrip is non-missing. Report the number of TRUE values.
```{r, echo =T}
#Part 5
length(as.character(Expends2002$CRPFilerid))
df$variable <- ifelse(grepl('N', as.character(Expends2002$CRPFilerid), ignore.case=T)==T, 1,0 
| as.numeric(Expends2002$Amount >500) & Expends2002$Descrip!="")
#df$variable #Variable indicating whether variable passes or fails above specifications
#Expends2002[,c("CRPFilerid", "Amount", "Descrip")]
```


    
    + EXTRA CREDIT: A variable that provides the most commonletter in the Descrip variable.
```{r, echo=T}
#Part 6
df$descrip <- gsub(" ", "", df$descrip, fixed = T)
names(table(unlist(strsplit(df$descrip, split =
""))))[as.vector(table(unlist(strsplit(df$descrip, split =
""))))==max(table(unlist(strsplit(df$descrip, split = ""))))]

```


**9. Write a script that subsets the data by state, and writes out a unique CSV file for each subset, where each file has a unique (and meaningful)name.**

```{r}
data <- by(Expends2002, Expends2002$State, function(x){Expends2002})
length(data)
#lapply(1:length(data), function(i) write.csv(data[[i]], file =
#paste0(names(data[i]), ".csv"), row.names = FALSE))

```



```