---
title: ""
output: pdf_document
header-includes:
- \newcommand{\benum}{\begin{enumerate}}
- \newcommand{\eenum}{\end{enumerate}}
- \newcommand{\beit}{\begin{itemize}}
- \newcommand{\eeit}{\end{itemize}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=F, results = 'asis')
```


\begin{center}
{\Large{\textbf{Problem Set 4}}} \\
\vspace{4 bp}
Due April 2, 10:00 AM (Before Class) \\
\end{center}

\section*{Instructions}

\benum
  \item The following questions should each be answered within an R script. Be sure to provide many comments in the script to facilitate grading. Undocumented code will not be graded. 
  \item Work on git. Fork the repository found at https://github.com/domlockett/PDS-PS3 and add your code, committing and pushing frequently. Use meaningful commit messages – these may affect your grade.
  \item You may work in teams, but each student should develop their own R script. To be clear, there should be no copy and paste. Each keystroke in the assignment should be your own.
  \item If you have any questions regarding the Problem Set, contact the TAs or use their office hours.
  \item For students new to programming, this may take a while. Get started.
  \item You will need to install \texttt{ggplot2} and \texttt{dplyr} to complete this dataset.
  
\eenum


\section*{Statistics}

\benum

 \item Load the following data: http://politicaldatascience.com/PDS/Datasets/GSS-data.csv. \\
 
 The variable \texttt{poleff11} asks participants to rate their level of agreement with the statement ``People like me don't have any say about what the government does'' (see the codebook for more information on all variables in this dataset at: http://politicaldatascience.com/PDS/Datasets/gss_codebook.csv). 

 \item Convert this variable into a numeric where higher values indicate higher levels of political efficacy (1- strongly agrees with the statement; 5- strongly disagrees with the statment)

```{r Q1.1, eval = T}
library(tidyverse)
data <- read_csv('http://politicaldatascience.com/PDS/Datasets/GSS-data.csv')
data$poleff <- recode(data$poleff11, "Strongly agree"= 1, "Agree" = 2, 'Neither agree nor disagree'=3, 'Disagree'=4, 'Strongly disagree'=5 )

```

 
\item What is the proportion of individuals from the entire sample who do not feel as though they have a say in the government? The proportion of those who do feel as though they have a say?

```{r Q1.2, eval = T}
sum(data$poleff < 3, na.rm=T)/sum(is.na(data$poleff)==F)
sum(data$poleff > 3, na.rm=T)/sum(is.na(data$poleff)==F)

```


\item Using a sample of \textbf{25} what is the average level of efficacy individuals feel on the one to five scale? At 100? 500?

```{r Q1.3 , eval =T}
mean(sample(data$poleff,size=25, replace =F), na.rm =T)
mean(sample(data$poleff,size=100, replace =F), na.rm =T)
mean(sample(data$poleff,size=500, replace =F), na.rm =T)
```

\item Create a variable called \texttt{trials\_25} where we pull a random sample of 25 from the \texttt{poleff11} data and calculate the mean 500 times . 


```{r Q1.4}
library(mosaic)

trials <- do(500)*mean(sample(data$poleff,size=25, replace =F), na.rm =T)

```

\item Now create a variable called \texttt{trials\_100} where we do 500 trials with a sample of 100 instead of 25. Draw a histogram of the sampling distribution fo the trials for the two trials you just conducted. Save these plots in your repository. 

```{r Q1.5}

trials <- do(1000)*mean(sample(data$poleff,size=100, replace =F), na.rm =T)

```

\item What notable difference occur when we use a larger sample size in our trials?



\section*{Bootstrapping}

\benum
  \setcounter{enumi}{1}

\item \textbf{Run the command \texttt{set.seed(25)}.} Create a new variable \texttt{eff_subset} which draws a sample of 300 observations from the original \texttt{poleff11} data. 

```{r Q2.1, eval=T}
set.seed(25)
eff_subset <- sample(data$poleff, size = 300, replace =F, na.rm=T)

```


\beit

\item Draw a sample of 25 from \texttt{eff_subset} and calculate the mean. Repeat this task 500 times and store the results in a variable named \texttt{trials}.

```{r Q2.2 , eval =T}
trials <- do(500)*mean(sample(eff_subset,size=25, replace =F), na.rm =T)

```

\item What is the mean of the 500 trials? 1000 trials? How do these values compare to the mean of the actual \texttt{poleff11} variable?

```{r Q2.3 , eval =T}
trials2 <- do(1000)*mean(sample(eff_subset,size=25, replace =F), na.rm =T)
mean(trials)
mean(trials2)
mean(data$poleff, na.rm=T)

```

\item Notice that with pulling random samples of only 25 observations from \textbf{a quarter of the data} we are able to accurately approximate the mean of our entire data. That is the logic of bootstrapping, with only a small sample (here our population is the entire dataset), one can make reasonable approximations about the population!


\eeit

\eenum


\section*{Supervised Learning}

 \item 

\beit
 \item 
 
```{r Q3.1, eval = F}

```

\item

```{r Q3.2, eval = F}

```

\item 

```{r Q3.3, eval = F}

```

\item 

```{r Q3.4, eval = T}
  
```

\item 

```{r Q3.5, eval = T}
```

\item 

```{r Q3.6, eval = T}


```

\item 

```{r Q3.7, eval = T}

```

\item 

```{r Q3.8, eval = T}

```


\item 


```{r Q3.9, eval = T}
```

\item 


```{r Q3.10, eval = T}

```


\eeit
\item 

```{r Q4 setup, eval= T}

```

\beit
\item 

```{r Q4.1}

```

\item 


```{r Q4.2}

```



\item 

```{r Q4.3}

```

\eeit



\eenum

